---
title: "sensorThings4R"
author: "Daniel Marsh-Hunn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
text-align: justify}
</style>

## Introduction

The [OGC SensorThings API](http://developers.sensorup.com/docs/) is a web standard for wireless sensor networks (WSN) and aims to provide an interoperable, unified framework for handling sensor data. The standard is based on elements from [OGC Sensor Web Enablement (SWE)](http://www.opengeospatial.org/ogc/markets-technologies/swe), but is designed to provide interconnectivity between Internet of Things (IoT) devices, data and applications over the web. By implementing SensorThings into WSNs, developers don't have to worry about heterogeneity of devices, data formats and web protocols. Web development using standards facilitates easier data accessibility, higher interoperability, and eventually lowers the costs of sensor and gateway providers.


## Motivation

The motivation for creating an R implementation of the SensorThings API emerged from a project in the scope of the author's master thesis. The project includes making an IoT sensor network in an agricultural environment SensorThings compliant. Since R is one of the leading programming environments for statistics and data analysis, but also includes software development functionalities, the idea materialised to write a package that uses the SensorThingsAPI to import data from the web. After importing the data, they can be used for further analysis. Furthermore, an interactive data visualisation tool for displaying the sensor locations and the station details was created using the functions within the package.


## How it works

SensorThings API relies on a set of classes that are interrelated using web URLs. Starting at a base URL (e.g. [https://toronto-bike-snapshot.sensorup.com/v1.0](https://toronto-bike-snapshot.sensorup.com/v1.0)), API data can be accessed by adding parameters to it. sensorThings4R makes use of the concept by defining a set of functions that compile SensorThings URLs and various parameters, use these in a http request, parse the resulting JSON data and store them in R dataframes for further usage.

## Installation

- Install with Git:
```{r, eval=FALSE}
devtools::install_github("danji90/sensorThings4R")
```
&nbsp;
&nbsp;

- Install manually with a .tar.gz file

&nbsp;
&nbsp;

## Functions

```{r, include=FALSE}
library(kableExtra)
```

To use the functions in sensorThings4R, first import the installed package:

```{r}
library(sensorThings4R)
```

#### Import JSON data into dataframe

Functions that request the data using the SensorThings base url, then parse and store the loaded JSON data in a dataframe. Depending on the function, the new dataframes are assigned the classes "locationObject", "foiObject" or "thingObject".

```{r}
# Load SensorThings Locations
loc = senseLocations("https://toronto-bike-snapshot.sensorup.com/v1.0")
class(loc)


# Load SensorThings FoIs
foi = senseFoI("https://toronto-bike-snapshot.sensorup.com/v1.0")
class(foi)


# Load SensorThings Things
s1 = senseThings("https://toronto-bike-snapshot.sensorup.com/v1.0")

class(s1)
```

&nbsp;
&nbsp;
&nbsp;


#### Processing for map visualisation

These functions convert "locationObject" and "foiObject" dataframes into new dataframes capable of being loaded to a leaflet map. The output dataframes have the added class "mapThing".

```{r}
level1 = senseLocations("http://example.sensorup.com/v1.0")
level2 = makeThingLocation(level1)

knitr::kable(level2) %>%
  kable_styling(bootstrap_options = "striped", font_size = 10) %>%
  scroll_box(width = "100%", height = "200px")

class(level2)

n = senseFoI("http://example.sensorup.com/v1.0")
u = makeThingFoI(n)

knitr::kable(tail(u)) %>%
  kable_styling(bootstrap_options = "striped", font_size = 10) %>%
  scroll_box(width = "100%", height = "200px")
```

&nbsp;
&nbsp;
&nbsp;

#### Mapping functions

The mapping functions load "mapThing" dataframes onto a leaflet map.

```{r, out.width='100%'}
x = senseLocations("http://example.sensorup.com/v1.0")
y = makeThingLocation(x)
mapThingLocations(y)

j = senseFoI("https://toronto-bike-snapshot.sensorup.com/v1.0")
k = makeThingFoI(j)
mapThingFoI(k)
```

&nbsp;
&nbsp;
&nbsp;

The methods only accept objects of the class they are designed for. An error message will indicate that the object does not have the right class and can't be mapped.

```{r,error=TRUE}
mapThingFoI(j)
```


&nbsp;
&nbsp;
&nbsp;

Express map functions combine the importing, processing and mapping functions to instantly create a map using an input SensorThings base url.

&nbsp;
&nbsp;
&nbsp;

```{r, out.width='100%'}
leaf1 = expressMapFoI("https://tasking-test.sensorup.com/v1.0")
leaf1
```

&nbsp;
&nbsp;
&nbsp;

#### ShinyThings

This function launches ShinyThings when called, a [Shiny application](https://shiny.rstudio.com) to display the locations in a sensor network and the station details. The UI provides simple user interaction with the app.

```{r, eval=FALSE}
shinyThings()
```

&nbsp;
&nbsp;
&nbsp;

